---
title: "Student Career Track"
author: "Muhamad Shaheedeazam"
date: "2023-12-20"
output: html_document
---

## Load necessary libraries

```{r loading packages}
library(readxl)
Student_Completion_Career_Track_Analysis <-
  read_excel("C:/Users/MuhamadShahee.H/Desktop/Azam/Portfolio/Student Completion Career Track Analysis.xlsx")
library(tidyverse)
library(here)
library(skimr)
library(janitor)
```

## Load and Explore the Data
#### View the dataset
```{r}
view(Student_Completion_Career_Track_Analysis)
```

#### Display the first few rows of the dataset
```{r}
head(Student_Completion_Career_Track_Analysis)
```

#### Check the structure of the dataset
```{r}
str(Student_Completion_Career_Track_Analysis)
```

#### Display column names
```{r}
colnames(Student_Completion_Career_Track_Analysis)
```

## Data Cleaning and Wrangling
#### Display summary statistics without charts
```{r}
skim_without_charts(Student_Completion_Career_Track_Analysis)
```

#### Glimpse at the dataset
```{r}
glimpse(Student_Completion_Career_Track_Analysis)
```

#### Select and filter columns
```{r}
Student_Completion_Career_Track_Analysis %>%
  select(track_name)
```

#### Select and filter columns, excluding 'track_name'
```{r}
Student_Completion_Career_Track_Analysis %>%
  select(-track_name)
```

#### Rename a column
```{r}
Student_Completion_Career_Track_Analysis %>%
  rename(role_name = track_name)
```

#### Clean column names
```{r}
clean_names(Student_Completion_Career_Track_Analysis)
```

## Sorting Data
#### Sort data by date_enrolled in ascending order
```{r}
Student_Completion_Career_Track_Analysis %>% arrange(date_enrolled)
```

#### Sort data by date_enrolled in descending order
```{r}
Student_Completion_Career_Track_Analysis %>% arrange(desc(date_enrolled))
```

#### Create a new dataset with sorted data
```{r}
Student_Completion_Career_Track_Analysis_V2 <- Student_Completion_Career_Track_Analysis %>%
  arrange(desc(date_enrolled))
View(Student_Completion_Career_Track_Analysis_V2)
```

## Grouping and Summarizing Data
#### Group data by 'days_for_completion' and calculate mean date_enrolled
```{r}
Student_Completion_Career_Track_Analysis %>%
  group_by(days_for_completion) %>%
  drop_na() %>%
  summarize(date_enrolled_2 = mean(date_enrolled))
```

#### Group data by 'track_name' and calculate maximum days_for_completion
```{r}
Student_Completion_Career_Track_Analysis %>%
  group_by(track_name) %>%
  drop_na() %>%
  summarize(completion_days = max(days_for_completion))
```

#### Group data by 'days_for_completion' and 'track_name', calculate mean date_enrolled
```{r}
Student_Completion_Career_Track_Analysis %>%
  group_by(days_for_completion, track_name) %>%
  drop_na() %>%
  summarize(date_enrolled_2 = mean(date_enrolled))
```

#### Group data by 'track_completed_status' and 'track_name', calculate max days_for_completion and mean date_enrolled
```{r}
Student_Completion_Career_Track_Analysis %>%
  group_by(track_completed_status, track_name) %>%
  drop_na() %>%
  summarize(max_CD = max(days_for_completion), mean_DE = mean(date_enrolled))
```

## Filtering Data
#### Filter data for completed tracks
```{r}
Student_Completion_Career_Track_Analysis %>%
  filter(track_completed_status == "Complete")
```

## Data Visualization
#### Create a bar chart showing completion status for each track
```{r}
ggplot(data = Student_Completion_Career_Track_Analysis) +
  geom_bar(mapping = aes(x=track_name, fill=track_completed_status)) +
  labs(title = "Student Completion Career Track", subtitle = "Career Track")
```

#### Create a bar chart showing completion status for each track with customized titles
```{r}
ggplot(data = Student_Completion_Career_Track_Analysis) +
  geom_bar(mapping = aes(x=track_name, fill=track_completed_status)) +
  labs(title = "Student Completion Career Track", subtitle = "Career Track")
```

#### Create a bar chart showing completion status for each track with customized titles
```{r}
ggplot(data = Student_Completion_Career_Track_Analysis) +
  geom_bar(mapping = aes(x=days_for_completion, fill=track_completed_status)) +
  labs(title = "Student Completion Career Track", subtitle = "Career Track")
```